---
title: "Analysis of victoria data"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 4
number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(DT)
```

# Read in data
## Keep data that passed the attention check
```{r}
textual_data <- read_csv("scales_text.csv", skip = 1)

pass_att_check <- read_csv("ulrik-data-text-12-03.csv")

data_to_use <- textual_data %>% 
  filter(`Please enter your prolific ID:` %in% pass_att_check$Q76)

```

## Use the dataset with textual responses
Our dataset looks like this:

```{r}
survey <- data_to_use %>% 
  janitor::clean_names() %>% 
  select(-x1) %>% 
  rename(ethnicity = how_would_you_best_describe_yourself_multiple_answers_possible_selected_choice,
         prolific_id = please_enter_your_prolific_id,
         devices_used = which_of_the_following_digital_devices_do_you_use) %>% 
  rename_with(~str_replace(.x, "in_the_past_week_on_average_approximately_how_much_time_per_day_have_you_spent_actively_using_your_", "daily_time_")) %>% 
  rename_with(~str_replace(.x, "think_about_how_you_use_your_qid64_choice_group_selected_choices_please_indicate_how_much_each_of_the_following_statements_reflects_your_experience_using_these_digital_devices_in_the_past_week_", "")) %>% 
  rename_with(~str_replace(.x, "below_are_a_number_of_statements_that_describe_ways_in_which_people_act_and_think_in_relation_to_their_digital_devices_please_indicate_how_much_you_agree_or_disagree_with_each_statement_in_relation_to_your_qid64_choice_group_selected_choices_", "")) %>% 
  select(prolific_id, everything())

survey %>% 
  datatable()

```


# Analysis
## Descriptives
### Ethnicity
```{r}
survey %>% 
  count(ethnicity) %>% 
  arrange(desc(n)) %>% 
  datatable()
```

### Device use
#### Devices they have
```{r}
survey %>% 
  count(devices_used) %>% 
  mutate(devices_used = fct_reorder(devices_used, n)) %>% 
  ggplot() +
    geom_col(aes(devices_used, n)) +
    coord_flip() + 
    labs(x = "", y = "")

props_use <- survey %>% 
  select(devices_used) %>% 
  mutate(devices_used = str_split(devices_used, ",")) %>% 
  unnest(cols = c(devices_used)) %>% 
  count(devices_used) %>% 
  mutate(ptc = (n / nrow(survey))*100)

props_use %>% 
  ggplot() +
    aes(x = reorder(devices_used, ptc), y = ptc) +
    geom_col() +
    coord_flip() +
    labs(x = "", y = "%")
```

#### Time spent
```{r}
usage_levels <- c(
  "More than 10 hours per day",
  "9-10 hours per day",
  "8-9 hours per day",
  "7-8 hours per day",
  "6-7 hours per day",
  "5-6 hours per day",
  "4-5 hours per day",
  "3-4 hours per day",
  "2-3 hours per day",
  "1-2 hours per day",
  "31-60 minutes per day",
  "10-30 minutes per day",
  "Less than 10 minutes per day"
)


time_spent <- survey %>% 
  select(prolific_id, starts_with("daily_time")) %>% 
  pivot_longer(cols = -prolific_id) %>% 
  rename(device = name, time = value) %>% 
  ungroup() %>% 
  mutate(time = factor(time, levels = usage_levels, ordered = TRUE))
  
time_spent %>% 
  filter(!is.na(time)) %>% 
  mutate(device = str_replace(device, "daily_time_", ""),
         device = str_replace(device, "_excluding_the_time_you_passively_wear_it", "")) %>% 
  ggplot() +
    geom_histogram(aes(fct_rev(time)), stat = "count") +
    coord_flip() +
    facet_wrap(~device, ncol = 7) +
    labs(title = "Daily time by device", subtitle = '"In the past week on average approximately how much time per day\nhave you spent actively using your..."', x = "")

```

### Browsers & tech savvy


## Scales
### BDSCS
#### Distribution of scores on individual items

Put the data in long format, in preparation for plotting it

```{r}
long_bdscs_scores <- survey %>% 
  select(prolific_id, in_the_past_week_i_was_good_at_resisting_temptation_on_my_digital_devices:in_the_past_week_i_often_acted_without_thinking_on_my_digital_devices) %>% 
  mutate(across(-prolific_id, ~case_when(
    . == "Not at all" ~ "1",
    . == "Very much" ~ "5",
    TRUE ~ .
  ))) %>% 
  pivot_longer(-prolific_id) %>% 
  rename(item = name, score = value)

long_bdscs_scores %>% 
  datatable()
```



#### Plot distribution of overall scores
We reverse score items that should be reverse-scored, so the plot it intuitive to read

```{r}
drama_colors <- c("darkgreen", "green", "orange", "red", "darkred")

non_reverse_items <- c(
  "in_the_past_week_i_was_good_at_resisting_temptation_on_my_digital_devices",
  "in_the_past_week_i_rejected_temptations_on_my_digital_devices",
  "based_on_the_past_week_people_would_say_that_i_have_iron_self_discipline_over_how_i_use_my_digital_devices"
)

bdscs_scores_reversed <- long_bdscs_scores %>% 
  mutate(scoring = if_else(item %in% non_reverse_items, "O", "R")) %>% 
  mutate(score = as.double(score),
         score = case_when(
    scoring == "R" & score == 1 ~ 5,
    scoring == "R" & score == 2 ~ 4,
    scoring == "R" & score == 4 ~ 2,
    scoring == "R" & score == 5 ~ 1,
    TRUE ~ score
  ))

# plot
bdscs_scores_reversed %>% 
  mutate(score = factor(score)) %>% 
  mutate(item = str_replace_all(item, "_", " ")) %>% 
  ggplot(aes(x = item, fill = fct_rev(score))) +
    geom_bar(position = "fill") +
    scale_fill_manual (values = drama_colors) +
    coord_flip() +
    scale_y_continuous(labels = scales::percent) +
    scale_x_discrete(labels = scales::label_wrap(70)) +
    labs(fill = "Self-control score\n(items reversed)", x = "", y = "",
         title = "Green means less impulsivity")

#ggsave("bdsc_distributions_victoria.png", width = 8, height = 5)
```


Plot a histogram of the scores
```{r}
summarised_scores <- bdscs_scores_reversed %>% 
  group_by(prolific_id) %>% 
  summarise(total_score = sum(score, na.rm = TRUE)) 

summarised_scores %>% 
  ggplot() +
    geom_histogram(aes(x = total_score))

summarised_scores %>% 
  summarise(mean = mean(total_score),
            sd = sd(total_score))
```



### UPPS
#### Distribution of scores on individual items, by dimension
Grab the scale information

```{r}
upps_items <- read_csv("upps_items.csv") %>% 
  janitor::clean_names() %>% 
  select(number_full, scoring_digital, dimension, current_item_for_digital_self_control) %>% 
  filter(!is.na(dimension)) %>% 
  rename(item = current_item_for_digital_self_control) %>% 
  mutate(item = janitor::make_clean_names(item),
         item = str_replace_all(item, "_", " "))
  

upps_items %>% 
  datatable()
```

Make the dataset long and join it with the item information, in preparation for plotting

```{r}
upps <- survey %>%
  select(prolific_id, i_have_a_reserved_and_cautious_attitude_towards_using_digital_devices:i_am_surprised_at_the_things_i_do_on_my_digital_devices_while_in_a_great_mood)


upps_scores_long <- upps %>% 
  pivot_longer(cols = -prolific_id) %>% 
  rename(item = name, score = value) %>% 
  mutate(item = str_replace_all(item, "_", " ")) %>% 
  left_join(upps_items) %>% 
  mutate(score = factor(score, levels = c("Strongly disagree", "Disagree", "Neither agree nor disagree", "Agree", "Strongly agree"), ordered = TRUE)) %>% 
  mutate(score_numeric = case_when(
    score == "Strongly disagree" ~ 5,
    score == "Disagree" ~ 4,
    score == "Neither agree nor disagree" ~ 3,
    score == "Agree" ~ 2,
    score == "Strongly agree" ~ 1
  )) %>% 
  # reverse numeric coding when item should be reverse scored
  mutate(score_numeric_reversed = case_when(
    scoring_digital == "R" & score_numeric == 1 ~ 5,
    scoring_digital == "R" & score_numeric == 2 ~ 4,
    scoring_digital == "R" & score_numeric == 4 ~ 2,
    scoring_digital == "R" & score_numeric == 5 ~ 1,
    TRUE ~ score_numeric)) %>% 
  mutate(score_numeric_reversed = factor(score_numeric_reversed))
  

```


```{r, fig.width=6, fig.height=4, out.width='100%', eval = FALSE}
# point plot for each dimension
draw_point_plot_for_a_dimension <- function(some_dimension){
  plot <- upps_scores_long %>% 
    filter(dimension == some_dimension) %>% 
    ggplot(aes(x = item, y = score)) +
      geom_boxplot() +
      geom_jitter(width = 0.3, height = 0.2, alpha = 0.2) +
      coord_flip() +
      labs(x = "", y = "", title = str_c("Dimension: ", some_dimension)) +
      scale_x_discrete(labels = scales::label_wrap(60))
  
  print(plot)
}

walk(upps_dimensions, draw_point_plot_for_a_dimension)
```

##### Plot the proportions of actual responses
Make a stacked bar chart for each dimension

```{r, fig.width=8, fig.height=6, out.width='100%'}
upps_dimensions <- c("premeditation",
                    "negative urgency",
                    "positive urgency",
                    "perseverance",
                    "sensation seeking")

draw_stacked_bar_chart_for_a_dimension <- function(some_dimension){
  plot <- upps_scores_long %>% 
    filter(dimension == some_dimension) %>% 
    ggplot(aes(x = item, fill = fct_rev(score))) +
      geom_bar(position = "fill") +
      coord_flip() +
      labs(x = "", y = "", title = str_c("Dimension: ", some_dimension), fill = "Score") +
      scale_x_discrete(labels = scales::label_wrap(60)) +
      scale_y_continuous(labels = scales::percent)
  
  print(plot)
}

walk(upps_dimensions, draw_stacked_bar_chart_for_a_dimension)


```

##### Plot the proportions with color coding for impulsivity
Make a stacked bar chart for each dimension

```{r, fig.width=8, fig.height=6, out.width='100%'}
upps_dimensions <- c("premeditation",
                    "negative urgency",
                    "positive urgency",
                    "perseverance",
                    "sensation seeking")

drama_colors <- c("darkred", "red", "orange", "green", "darkgreen")

draw_stacked_bar_chart_for_a_dimension <- function(some_dimension){
  plot <- upps_scores_long %>% 
    filter(dimension == some_dimension) %>% 
    ggplot(aes(x = item, fill = fct_rev(score_numeric_reversed))) +
      geom_bar(position = "fill") +
      scale_fill_manual (values = drama_colors) +
      coord_flip() +
      labs(x = "", y = "", title = str_c("Dimension: ", some_dimension, "\n(Red means more impulse)"), fill = "Score") +
      scale_x_discrete(labels = scales::label_wrap(60)) +
      scale_y_continuous(labels = scales::percent)
  
  print(plot)
}

walk(upps_dimensions, draw_stacked_bar_chart_for_a_dimension)


```


<!-- Get the numbers -->
```{r, eval=FALSE}
upps_scores_long %>% 
  count(item, score, dimension) %>%
  arrange(dimension) %>% 
  mutate(ptc = (n/nrow(survey))*100,
         ptc = round(ptc, 1)) %>% 
  pivot_wider(names_from = score, values_from = c(n, ptc)) %>% 
  datatable()
```

#### Distribution of mean scores on each dimension
Note that the original scale ["uses a 1 (agree strongly) to 4 (disagree strongly) response format"](http://www.impulsivity.org/measurement/UPPS-P_59_item.pdf).
In keeping with the original scale, we code such that "Strongly disagree" is a 5 and "Strongly agree" is a 1.


We first do the numerical coding, then calculate the mean score for each participant

```{r}
dimension_scores_by_participant <- upps_scores_long %>% 
  group_by(prolific_id, dimension) %>% 
  summarise(scale_score = mean(as.double(score_numeric_reversed), na.rm = TRUE)) %>% 
  ungroup()

```

Now let's plot it

```{r}
dimension_scores_by_participant %>% 
  mutate(dimension = factor(dimension, levels = upps_dimensions)) %>% 
  ggplot(aes(x = dimension, y = scale_score)) +
      geom_boxplot() +
      geom_jitter(width = 0.2, height = 0, alpha = 0.3) +
      labs(x = "", y = "", title = "Mean impulsivity score by dimension")
```

