---
title: "Factor analysis"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(corrr)
library(psych)
library(DT)
```

# Factor analysis
Following instructions from here: http://www2.hawaii.edu/~georgeha/Handouts/meas/Exercises/_book/efa.html


## Read in data
### Keep data that passed the attention check
```{r}
pass_att_check <- read_csv("ulrik-data-text-12-03.csv")

textual_data <- read_csv("scales_text.csv", skip = 1) %>% 
  filter(`Please enter your prolific ID:` %in% pass_att_check$Q76)

# do sensible column naming
data_to_use <- textual_data %>% 
  janitor::clean_names() %>% 
  select(-x1) %>% 
  rename(prolific_id = please_enter_your_prolific_id) %>% 
  rename_with(~str_replace(.x, "think_about_how_you_use_your_qid64_choice_group_selected_choices_please_indicate_how_much_each_of_the_following_statements_reflects_your_experience_using_these_digital_devices_in_the_past_week_", "")) %>% 
  rename_with(~str_replace(.x, "below_are_a_number_of_statements_that_describe_ways_in_which_people_act_and_think_in_relation_to_their_digital_devices_please_indicate_how_much_you_agree_or_disagree_with_each_statement_in_relation_to_your_qid64_choice_group_selected_choices_", ""))

```

### Join the UPPS-P columns with the item info

```{r}
upps_items <- read_csv("upps_items.csv") %>% 
  janitor::clean_names() %>% 
  select(number_full, scoring_digital, dimension, current_item_for_digital_self_control) %>% 
  filter(!is.na(dimension)) %>% 
  rename(item = current_item_for_digital_self_control) %>% 
  mutate(item = janitor::make_clean_names(item),
         item = str_replace_all(item, "_", " "))

upps_scores_long <- data_to_use %>%
  select(prolific_id, i_have_a_reserved_and_cautious_attitude_towards_using_digital_devices:i_am_surprised_at_the_things_i_do_on_my_digital_devices_while_in_a_great_mood) %>% 
  pivot_longer(cols = -prolific_id) %>% 
  rename(item = name, score = value) %>% 
  mutate(item = str_replace_all(item, "_", " ")) %>% 
  left_join(upps_items) %>% 
  mutate(score_numeric = case_when(
    score == "Strongly disagree" ~ 5,
    score == "Disagree" ~ 4,
    score == "Neither agree nor disagree" ~ 3,
    score == "Agree" ~ 2,
    score == "Strongly agree" ~ 1
  )) %>% 
  mutate(item_label = str_c(number_full, "_", dimension))
  
upps_scores_wide <- upps_scores_long %>% 
  select(prolific_id, item_label, score_numeric) %>% 
  mutate(item_label = str_replace(item_label, "premeditation", "premed"),
         item_label = str_replace(item_label, "negative urgency", "neg_urg"),
         item_label = str_replace(item_label, "positive urgency", "pos_urg"),
         item_label = str_replace(item_label, "sensation seeking", "sen_seek"),
         item_label = str_replace(item_label, "perseverance", "persev")) %>% 
  pivot_wider(names_from = item_label,
              values_from = score_numeric)
```

So the data set we'll use for factor analysis looks like this:

(1: we're looking at the numerically coded scores, where 5 means 'strongly disagree'

2: we *have not* reverse scored items that should be reversed
)

```{r}
upps_scores_wide %>% 
  datatable()
```

## Describe the data
### Summary statistics
```{r}
upps_scores_only <- upps_scores_wide %>% select(-prolific_id)

psych::describe(upps_scores_only) %>% 
  datatable()
```

### Correlation matrix
```{r}
upps_correlations <- upps_scores_only %>% 
  correlate() %>% 
  rearrange() %>%  # rearrange by correlations
  shave() # Shave off the upper triangle for a clean result

fashion(upps_correlations) %>% 
  datatable()

rplot(upps_correlations)
```

### KMO factor adequacy & Bartlett's sphericity test
```{r}
KMO(upps_scores_only)

cortest.bartlett(upps_scores_only)
```


### Parallel analysis
```{r}
library(nFactors)

n_p  <- sum(complete.cases(upps_scores_only)) # The number of participants in our data

n_nu <- ncol(upps_scores_only) # The number of variables in our data

reduced_eig <- eigenComputes(as.data.frame(upps_scores_only), model = "factors", use = "complete")

n_factors  <- length(reduced_eig)

paral <- parallel(subject = n_p,  
                      var = n_nu, 
                      rep = 100,
                 quantile = .95, 
                   model  = "factors")

parallel_ana <- data.frame(Nfactor = 1:n_factors,
                           reduced_eig,
                           RandEigM = paral$eigen$mevpea,
                           RandEig95= paral$eigen$qevpea)
parallel_ana <- round(parallel_ana, 3)

parallel_ana
```

### Scree plot
#### reduced correlation matrix

```{r}
scree <- data.frame(factor_n = as.factor(1:n_factors), 
                  eigenvalue = reduced_eig)

ggplot(scree, aes(x = factor_n, y = eigenvalue, group = 1)) + 
  geom_point() + geom_line() +
  xlab("Number of factors") +
  labs( title = "Scree Plot", 
        subtitle = "(Based on the reduced correlation matrix)")
```

#### Unreduced correlation matrix
```{r}
fafitfree <- fa(upps_scores_only, nfactors = ncol(upps_scores_only), rotate = "none")

n_factors <- length(fafitfree$e.values)

scree     <- data.frame(
               Factor_n =  as.factor(1:n_factors), 
             Eigenvalue = fafitfree$e.values)

ggplot(scree, aes(x = Factor_n, y = Eigenvalue, group = 1)) + 
  geom_point() + geom_line() +
  xlab("Number of factors") +
  ylab("Initial eigenvalue") +
  labs( title = "Scree Plot", 
     subtitle = "(Based on the unreduced correlation matrix)")
```



## Factor analysis
### Using raw data, and assume data are continuous
```{r}
fafit <- fa(upps_scores_only, nfactors = 4, fm = "ml", rotate = "promax")

n_factors <- length(fafit$e.values)

print(fafit, cut = .32, sort = TRUE, digits = 3)
```

### Assume ordinal data
